```{r}
D1 <- read.csv("~/Documents/Rstudio/assignment 4/Assistments-confidence.csv", header = TRUE) 
```

```{r}
  D1$id <- NULL
```

## Create a correlation matrix of the relationships between the variables, including correlation coefficients for each pair of variables/features.

```{r}
library(corrplot)
COR <- cor(D1)

corrplot(COR, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")
```

```{r}
D2 <- read.csv("~/Documents/Rstudio/assignment 4/Assistments-confidence.csv", header = TRUE) 
D2$mean_correct <- NULL
D2 <- scale(D2, center = TRUE)
pca <- prcomp(D2, scale = TRUE)
```

```{r}
pca$sdev

#To convert this into variance accounted for we can square it, these numbers are proportional to the eigenvalue

pca$sdev^2

#A summary of our pca will give us the proportion of variance accounted for by each component

summary(pca)

#We can look at this to get an idea of which components we should keep and which we should drop

plot(pca, type = "lines")
```

## Think about which components you would drop and make a decision
## I will drop 5, 6, 7 because variances in 1-4 are big and there is a slope since 4.



```{r}
D3 <- as.data.frame(pca$x)

#Attach the variable "mean_correct" from your original data frame to D3.

D4 <- cbind(D3, as.data.frame(D1$mean_correct))

#Now re-run your scatterplots and correlations between the transformed data and mean_correct. If you had dropped some components would you have lost important infomation about mean_correct?


COR2 <- cor(D4)

corrplot(COR2, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")
```

## I think PC7 and PC5 stil can be dropped but if I drop PC6, I could lost a lot of imformation.

```{r}
pca$rotation

#Examine the eigenvectors, notice that they are a little difficult to interpret. It is much easier to make sense of them if we make them proportional within each component

loadings <- abs(pca$rotation) #abs() will make all eigenvectors positive
  
sweep(loadings, 2, colSums(loadings), "/") #sweep() computes each row as a proportion of the column. (There must be a way to do this with dplyr()?)

#Now examine your components and try to come up with substantive descriptions of what some might represent?

# PC1 might represent that prior problem counts and mean hint are highly related, which may suggest that the more questions students try to answer, the more hints they need.
# PC2 might mean that as a result of the larger proportion of prior_prob_count and mean attempt.
# PC3 the relationship between prior_percent_correct and problems_attempted may indicate that the diligence of students.

#You can generate a biplot to help you, though these can be a bit confusing. They plot the transformed data by the first two components. Therefore, the axes represent the direction of maximum variance. Then mapped onto this point cloud are the original directions of the variables, depicted as red arrows. It is supposed to provide a visualization of which variables "go together". Variables that possibly represent the same underlying construct point in the same direction.  

biplot(pca)

#Calculate values for each student that represent these your composite variables and then create a new correlogram showing their relationship to mean_correct.

```


# Part III  
## Also in this repository is a data set and codebook from Rod Martin, Patricia Puhlik-Doris, Gwen Larsen, Jeanette Gray, Kelly Weir at the University of Western Ontario about people's sense of humor. Can you perform a PCA on this data?

```{r}
DF <- read.csv("~/Documents/Rstudio/assignment 4/humor_data.csv", header = TRUE) 
humor <- DF[1:32]
```

```{r}
COR_humor <- cor(humor)

corrplot(COR_humor, order="AOE", method="circle", tl.pos="lt", type="upper",        
tl.col="black", tl.cex=0.6, tl.srt=45, 
        addCoef.col="black", addCoefasPercent = TRUE,
        sig.level=0.50, insig = "blank")
```

```{r}
H1 <- scale(humor, center = TRUE)
pca2 <- prcomp(H1, scale = TRUE)

pca2$sdev
pca2$sdev^2
summary(pca2)
plot(pca2, type = "lines")

pca2$rotation

loadings2 <- abs(pca2$rotation) 
  
sweep(loadings2, 2, colSums(loadings2), "/")


biplot(pca)
```


